\documentclass[a4paper]{article}
\topmargin-2.0cm

%\usepackage{lineno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amscd}
\usepackage{amsfonts}
\usepackage{graphicx}%
\usepackage{fancyhdr}
\usepackage[hyphenbreaks]{breakurl}
\usepackage[hyphens]{url}

\newcommand{\ttbarV}{\ensuremath{t\bar{t}V}}
\newcommand{\ttbarW}{\ensuremath{t\bar{t}W}}
\newcommand{\ttbarZ}{\ensuremath{t\bar{t}Z}}
\newcommand{\ifb}{\ensuremath{\mathrm{fb}^{-1}}}

\setlength{\topmargin}{-15mm}
\setlength{\textwidth}{7in}
\setlength{\oddsidemargin}{-8mm}
\setlength{\textheight}{9in}
\setlength{\footskip}{1in}


%\title{\vspace{-3cm} Research Statement}
%\author{Othmane Rifki}%\\
%University of Oklahoma\\
%E-mail: \texttt{othmane.rifki@cern.ch}}
%\date{December 2016}


\begin{document}

\thispagestyle{fancy}\lhead{\large{Research Statement}} \rhead{\large{May 22, 2020}}
\chead{{\large{\bf Othmane Rifki}}} \lfoot{} \rfoot{\bf \thepage} \cfoot{}

%\linenumbers
\fontsize{11}{14}
\selectfont
%\maketitle


My research interests lie in exploring the Standard Model (SM) of particle physics and at the same time addressing questions not answered by the SM, such as the nature of dark matter (DM). Answering these questions requires improving detector capabilities and exploring new analysis ideas to enhance the precision of our measurements and to maximize the discovery potential of the ATLAS detector. The majority of searches for beyond the SM (BSM) physics in ATLAS focuses on signatures with promptly decaying new particles, leaving a wide uncovered parameter space of particles with detectable displacements between their production and decay points. At Columbia, I would like to develop
novel techniques to search for long-lived particles and to make important contributions to the readout electronics of the next-generation ATLAS liquid argon (LAr) calorimeter system, benefiting from and contributing to the expertise of the group in these areas.

\bigskip

So far, my research activities have spanned several areas in physics analysis and detector development. As a DESY fellow, I led a search for DM in invisible Higgs boson decays, setting the strongest existing limit to date, and built a module loading station for the new all-silicon Inner Tracker (ITk) upgrade for the High-Luminosity LHC (HL-LHC). During my PhD, I performed the first searches for supersymmetry with early Run-2 data in leptonic final states after nearly doubling the center-of-mass energy to 13 TeV. To cope with the increased data readout rates in Run-2, I migrated the functionality of the Region of Interest Builder (RoIB), the interface that seeds the processing of every event in the high level trigger farm (HLT), from a VMEbus system to a PCIe card running in a commodity-computer. This new system has been the baseline for data taking in ATLAS since the start of 2016.

\bigskip

At DESY, I have been deeply involved in the upgrade of the ITk strip detector\cite{ITKstrips}. I led the project to successfully develop a fully automated assembly system that places strip modules onto a petal, the basic unit that forms the ITk end-cap disks. This system consists of mounting modules with less than 50 $\mu m$ lateral accuracy onto a petal core using a robotic gantry as a pick-and-place machine with advanced vision for identification of the sensor fiducial markers. I oversaw the purchasing of the equipment needed to build the station, totaling over $\$200$k, and exchanged with industry providers to test new solutions. I collaborated closely with other loading sites to standardize the module mounting procedure and adapt a common interface to streamline production. In February 2020, I successfully presented the first fully loaded petal built at DESY using this automated procedure to the module loading Final Design Review, showing the petal assembly meets the desired specifications with high repeatability and increased capacity to make up for possible production delays.

\bigskip

%The ITk will extend the angular coverage of the ATLAS inner detector which will help the physics of forward objects, such as the case of Higgs boson production in vector boson fusion (VBF). Using this VBF production, I coordinated the most sensitive search that drives the ATLAS direct probe of Higgs boson decays to invisible particles
The rate of invisible Higgs boson decays may be significantly enhanced beyond its SM value of $\sim0.1\%$ if the Higgs boson couples to DM particles.
The latter is a reasonable assumption given that the Higgs boson couples to massive particles and thus may act as a portal to DM, and the large uncertainties on its decay rates allow for additional decay to BSM particles. To test this hypothesis, I  via the VBF production mechanism \cite{vbfMET_CONF2020}. The VBF signature is a powerful probe for new physics as it exhibits a striking signature of two forward jets that have large invariant mass. This signature can be exploited to reject large SM backgrounds while retaining signal. I established a collaboration between theorists and the ATLAS analysis team to improve filtering algorithms used in the Monte Carlo generator, which resulted in the reduction of the largest uncertainty in the previous publication by 35\%. This work also enhanced the theoretical accuracy and precision of the simulation, which allowed us to better control the main $Z\left(\nu\nu\right)$+jets background.
I also oversaw the optimization of the event selection and analysis object definition to increase the Higgs boson signal acceptance by 50\% and reduce the background in the signal regions by 31\%. Through these improvements, my team and I significantly increased the sensitivity to this fundamental probe of dark matter by 54\%. Over the coming months, I would like to continue leading the effort of this search to complete a
legacy Run-2 paper and contribute to the combination effort of my analysis with other direct and indirect searches for invisible Higgs boson decays.

\bigskip

Looking ahead, the next years are of great importance to prepare for Run-3 and HL-LHC in terms of new analysis ideas. As there has been no definitive sign of new physics,
it is important to fully exploit the LHC by looking in experimentally hard-to-reach locations that have not yet been probed.
One potential avenue that I would like to explore which offers excellent sensitivity for well motivated scenarios is the search for long-lived particles \cite{alimena2019searching}.
Analogous to the SM that has particles spanning a wide range of lifetimes, models beyond the SM (BSM) also predict new particles
with a variety of lifetimes, many of which can lead to macroscopic displacements from the primary interaction
such as the case of .
While the long-lived particles in the SM, such as b-hadrons, have well-understood experimental signatures, BSM long-lived particles
offer unusual signatures that may be rejected by standard reconstruction algorithms.
Currently, tracks corresponding to these hypothetical particles are reconstructed with a dedicated tracking configuration that only runs
on a filtered part of the main physics dataset due to significantly larger computing cost and data size. This cost can be significantly reduced
by employing novel machine learning techniques to make a better track seed selection, thus including long-lived particle tracking
into the standard reconstruction algorithms. This would have a considerable impact on many other downstream algorithms, such as tagging of different
physics objects that can be used in long-lived particle searches, and will therefore expand the search program for these particles in ATLAS.
With my experience in optimizing software performance, I want to drive the long-lived particle tracking capability even further to expand
the phase space probed by the LHC.

\bigskip

Another avenue through which to improve these searches is to exploit the utility of prompt associated objects.
I am interested, for instance, in continuing to explore the Higgs sector through the exotic decay of the Higgs boson to neutral long-lived particles
via associated production with a vector boson. Columbia is already involved in a similar analysis looking for the Higgs boson
decaying to long-lived SUSY particles that decay into a photon, which does not point back to the primary vertex, and an invisible particle.
If these long-lived SUSY particles decay in the inner detector, the analysis uses the pointing and timing resolution of the LAr electromagnetic calorimeter to distinguish between prompt and displaced photons.
It would be useful to investigate the use of tracking by reconstructing the converted photon vertex and its trajectory to determine
its direction with respect to the primary vertex. The converted photons will have a lower statistics but an improved precision that
can potentially bring additional sensitivity to displaced photon searches.
I would like to explore similar non-standard ideas that capitalize on the ATLAS detector capabilities to develop new search strategies for Run-3.


\bigskip

%During my PhD, I sought an opportunity to work on the trigger and data acquisition (TDAQ) system
%of ATLAS at Argonne National Laboratory (ANL). I received an ATLAS Support Center (ASC) fellowship to prototype and develop
%an evolved Region of Interest Builder (RoIB), the system that collects the regions of interest
%found by the hardware trigger of ATLAS and forwards them to the software High Level trigger (HLT).
%I successfully installed and commissioned the system I built with the ANL colleagues which is currently used as baseline for data taking in ATLAS %since the start of 2016. In addition to the expert knowledge I acquired in TDAQ, I also got an exposure to ATLAS operations where I had
%to interact with both trigger and DAQ experts to understand the requirements of the different subsystems of TDAQ  while integrating the RoIB.

%My work in the strip ITk community has prepared me well to join the ANL group to work on the assembly and testing of the silicon pixel modules.
%I have already performed some of the tasks that will be required at ANL as part of the module loading procedure; prior to loading the modules on the %local support structure, I verified that modules are qualified for assembly by visually inspecting the modules and
%facilitating the wire-bonds repairs prior to electrically testing the modules.
%With this experience, I want to learn from and contribute to ANL's work to prepare for production of the ITk pixel modules expected to ramp up during %the upcoming years. I want to expand my detector experience to building and testing modules, as well as characterizing them in test beams.

%I would like to expand my experience working with the tracker to working with the calorimeter.
%based on my phd experience in readout.

I hope to leverage my expertise in BSM searches, detector development, and data acquisition systems to continue searching for new physics and
prepare for the production and installation of the next generation LAr readout electronics.
I also aim to assume more responsibilities within ATLAS. The work I have done
during my DESY fellowship as well as in my PhD have prepared me well to take on new challenges.
Columbia plays a central role in both searches for new physics and hardware upgrades, and I would be delighted to work,
learn, and grow as part of the group.


% physics and LAr upgrade communities

% https://atlas.web.cern.ch/Atlas/GROUPS/PHYSICS/PAPERS/EXOT-2017-25


\Urlmuskip=0mu plus 1mu\relax
{%\small
\fontsize{11}{14}
\selectfont
\bibliography{research_statement}}{}
\bibliographystyle{unsrt}

\end{document}
